---
title: "Rapport - Séries Temporelles"
output:
  html_document:
    df_print: paged
---

#### Nom et Prénom des étudiants du groupe :

```
- REAL : Philippe : 
```

#### Instructions

* Vous rendrez un compte rendu écrit sous format HTML en faisant votre TP directement dans ce fichier R Markdown.
* Utiliser le bouton **Knit HTML** dans RStudio pour compiler le fichier.
* Vous écrirez vos programmes R complets dans des blocs de code et vos commentaires et explications dans des paragraphes de texte. 
* Vous pouvez utiliser la syntaxe LaTeX pour écrire des formules mathématiques dans les parties de texte.
* N'hésitez pas à consulter la documentation relative à RMarkdown -> bouton **?** ou sur http://rmarkdown.rstudio.com/
* Pensez à écrire vos noms et prénoms en haut du document.
* Vous enverrez votre fichier au plus tard le **//2015** au soir à mon adresse jonathan.el-methni@parisdescartes.fr  


```{r eval=FALSE, message=FALSE, warning=FALSE}

#for manipulate data (transform to dataframe)
install.packages("tidyverse")
install.packages("tibble")
install.packages("sm")

install.packages("KernSmooth")
install.packages("np")
install.packages("stats")
install.packages("ggplot2")
install.packages("kedd")
```


```{r} 
install.packages("forecast")
install.packages("TSA")
#install.packages("aTSA")
install.packages("tseries")
install.packages("TTR")
install.packages("gridExtra")
```

## Load libraries 
```{r r_utile, echo=FALSE, message=FALSE, warning=FALSE}

rm(list=ls())

#chargement des librairies utilses
library(tibble)
library(ggplot2)
library(KernSmooth)
library(stats)
library(np)
library(kedd)
library(tidyverse)

library(forecast)
library(TSA)
#library(aTSA)
library(tseries)
library(TTR)
library(gridExtra)
require(graphics)
```


## 2ème Exemple : CAC-40

#### Chargement des données
Les données ont été récupérées sur le site Yahoo Finance

https://finance.yahoo.com/quote/%5EFCHI/history?period1=879289200&period2=1573513200&interval=1d&filter=history&frequency=1d

https://www.kaggle.com/kp4920/s-p-500-stock-data-time-series-analysis/code


### 1.0 Lecture des données et premières analyses

* Read Data
```{r d1_summary}
dataCAC40_raw <-read.table("Data_CAC40_1997-2019.csv", sep=",", dec=".",header=T, na.strings = "null")
```

* Data overview
```{r}
summary(dataCAC40_raw)

```

* Data cleaning
The data has some missing values, which we will replace with zero. Also, 'Date' feature is listed as factor, we will convert that to 'Date' structure.

```{r}

dataCAC40<-dataCAC40_raw[which(dataCAC40_raw$Open != "NA"),]
#dataCAC40<-dataCAC40_raw
#dataCAC40[is.na(dataCAC40)] <- 0


#nbNull<-which(dataCAC40[,2]=="null")
FrameCAC40 <- as.data.frame(dataCAC40)

FrameCAC40[['Date']] <- as.Date(FrameCAC40[['Date']], format='%Y-%m-%d')  
#options(digits=10)
#FrameCAC40[['Open']] <- as.double(FrameCAC40[['Open']])

FrameCAC40[,2] <- as.numeric(as.character(FrameCAC40[,2]))
FrameCAC40[,3] <- as.numeric(as.character(FrameCAC40[,3]))
FrameCAC40[,4] <- as.numeric(as.character(FrameCAC40[,4]))
FrameCAC40[,5] <- as.numeric(as.character(FrameCAC40[,5]))
FrameCAC40[,6] <- as.numeric(as.character(FrameCAC40[,6]))

head(FrameCAC40)
summary(FrameCAC40)

```

```{r} 

library(ggplot2)
options(repr.plot.width=12, repr.plot.height=12) 
p1 = ggplot(FrameCAC40, aes(Open)) + geom_histogram(bins = 50, aes(y = ..density..), col = "red", fill = "red", alpha = 0.3) + geom_density()# + xlim(c(0, 1000))
p2 = ggplot(FrameCAC40, aes(Close)) + geom_histogram(bins = 50, aes(y = ..density..), col = "red", fill = "red", alpha = 0.3) + geom_density()# + xlim(c(0, 1000))

#grid.arrange(p1,p2, nrow=1,ncol=2)
grid.arrange(p1, nrow=1,ncol=1)
```



```{r fig.width=10}
#CREATE DAILY OBJECT
inds<-seq(as.Date("1997-11-12"	),as.Date("2019-11-12"), by ="day")
#crate function tahatgenerate time
create_ts <- function(col_idx){
  i_ts<-as.numeric(FrameCAC40[,col_idx]) %>%
    ts(start = c(1997,as.numeric(format(inds[1],"%j"))), frequency =365.5 )
  return (i_ts)
}

i_ts = create_ts(colnames(FrameCAC40) == "Open")
plot(i_ts,xlab="Time",ylab="Close",main="Time serie - CAC40",col="red")

```

```{r} 
library(zoo)
library(xts)
library(lubridate)
#○library( TStudio)

xtsCAC40 <- xts(x = FrameCAC40, order.by = FrameCAC40$Date)

library(readxl)
start <- as.Date('1997-11-12')
end <- as.Date('2019-11-12')
tsCAC40<-ts(xtsCAC40, start=start, frequency = 365)
#tsCAC40<-as.timeSeries.xts(dataCAC40,frequency = d,start ='1997-11-12')
class(tsCAC40)

frequency(tsCAC40)

#tsCAC40_Close=ts(tsCAC40(Open),start=c(1997,11,12))
```




### Time Series Analysis

Before we start with the time series analysis, lets go through the theory in brief.

What is AutoRegressive or AR model:

Autoregressive (AR) models are models where the value of variable in one period is related to the values in the previous period.
AR(p) is a Autoregressive model with p lags.
  
What is Moving Average or MA model:

Moving average (MA) model accounts for the possibility of a relationship between a variable and the residual from the previous period.
MA(q) is a Moving Average model with q lags.

What is ARMA model:
Autoregressive moving average model combines both p auto regressive terms and q Moving average terms, also called ARMA(p,q)



* Transfom data to time serie
```{r fig.width=15}
tsCAC40_Close=ts(as.vector(FrameCAC40[,5]),start=as.Date("1997-11-12"),frequency=365)

plot(tsCAC40_Close,xlab="Time",ylab="Close",main="Time serie - CAC40",col="red")

#monthplot(tsCAC40_Close)

```


* Transfom data to time serie
```{r fig.width=15}
FrameCAC40_2012 <- FrameCAC40[FrameCAC40[,1]> as.Date("2012-11-11"),]

tsCAC40_2012_Close=ts(as.vector(FrameCAC40_2012[,5]),start=2012,frequency=365)
```

## Plotting Time Series - Close

There are 5 time series in the data provided - (High, Low, Open, Close, Volume). We will look at the High values first.

```{r fig.width=15}

plot(tsCAC40_Close,xlab="Time",ylab="Close",main="Time serie - CAC40 - 1997/2019")#,col="red")

plot(tsCAC40_2012_Close,xlab="Time",ylab="Close",main="Time serie - CAC40 - 2012/2019")#,col="red")

#monthplot(tsCAC40_Close)

```


```{r fig.width=25}

monthplot(tsCAC40_2012_Close)
```


### Stationarity
What is stationary time series?
A stationary process has a mean and variance that do not change overtime and the process does not have trend.
The above time series does not look stationary.
To confirm that we will use "Dickey-Fuller test" to determine stationarity. 
  
Dickey-Fuller test for variable
 
```{r} 

adf.test(as.vector(FrameCAC40[,5]),alternative="stationary",k=0)
adf.test(as.vector(FrameCAC40_2012[,5]),alternative="stationary",k=0)

```
 
### Decomposing Time Series
Decomposing a time series involves separating the time series into trend and irregular components.

```{r}
i_tscomponents <- decompose(tsCAC40_Close)
plot(i_tscomponents)#, col = "red")

i_tscomponents_2012 <- decompose(tsCAC40_2012_Close)
plot(i_tscomponents_2012)#, col = "red")
```


### Differencing a Time Series
Differencing is a common solution used to stationarize the variable.
We will perform differencing using R function diff.


```{r fig.width=15}

i_tsdiff<-diff(tsCAC40_Close,differences = 1)
plot(i_tsdiff)#,col="red")

i_tsdiff_2012<-diff(tsCAC40_2012_Close,differences = 1)
plot(i_tsdiff_2012)#,col="red")

```

The time series (above) appears to be stationary.



## Selecting a Candidate ARIMA Model
The next step is to select appropriate ARIMA model, which means finding the most appropriate values of p and q for an ARIMA(p,d,q) model. You usually need to examine the correlogram and partial correlogram of the stationary time series for this.
To plot a correlogram and partial correlogram, we can use the acf() and pacf() functions in R, respectively.

```{r}
acf(i_tsdiff, lag.max=60)             # plot a correlogram
acf(i_tsdiff, lag.max=60, plot=FALSE) # get the autocorrelation values

```
```{r}

pacf(i_tsdiff, lag.max=60)             # plot a partial correlogram
pacf(i_tsdiff, lag.max=60, plot=FALSE) # get the partial autocorrelation values

```


```{r}

acf(i_tsdiff_2012, lag.max=60)             # plot a correlogram
acf(i_tsdiff_2012, lag.max=60, plot=FALSE) # get the autocorrelation values
```

```{r}

pacf(i_tsdiff_2012, lag.max=60)             # plot a partial correlogram
pacf(i_tsdiff_2012, lag.max=60, plot=FALSE) # get the partial autocorrelation values

```

Now, we could compare the sample ACF and PACF to those of various theoretical ARMA models. Use properties of ACF & PACF as a guide to estimate plausible models and select appropriate p, q and d. Alternative to this is discussed next.

## Fitting an ARIMA Model
  R provides a function auto.arima, which returns best ARIMA model according to either AIC, AICc or BIC value. The function conducts a search over possible model within the order constraints provided.
  
```{r}
i_tsarima <- auto.arima(tsCAC40_Close, max.p = 3, max.q = 3, max.d = 3)
i_tsarima
```

```{r}
i_tsarima_2012 <- auto.arima(tsCAC40_2012_Close, max.p = 3, max.q = 3, max.d = 3)
i_tsarima_2012
```


## Forecasting using an ARIMA Model
```{r}
help("forecast")

i_tsforecasts <- forecast(i_tsarima, h = 60)
plot(i_tsforecasts, col = "red")
```

```{r}
plot.ts(i_tsforecasts$residuals)            # make time plot of forecast errors
ggplot(data.frame(residuals = i_tsforecasts$residuals), aes(residuals)) + geom_histogram(bins = 50, aes(y = ..density..), col = "red", fill = "red", alpha = 0.3) + geom_density()# make a histogram
```

The forecast errors seem to be normally distributed with mean zero and constant variance, the ARIMA model does seem to provide an adequate predictive model


Here we looked at how to best fit ARIMA model to univariate time series.
Next thing that I'll work on is Multivariate Time Series Forecasting using neural net.



```{r fig.width=15}



```

#### 1. Vous pouvez mettre en avant votre texte de différentes façons

*"Essentially, all models are wrong, but some are useful" G.Box*

**"Essentially, all models are wrong, but some are useful" G.Box**

> "Essentially, all models are wrong, but some are useful" G.Box

```
"Essentially, all models are wrong, but some are useful" G.Box
```

Trait horizontal de délimitation 

**********

Différentes tailles avec #

# Test
## Test
### Test
#### Test

#### 2. Vous pouvez écrire vos codes R. Les résultats s'afficheront à la suite.
 
```{r} 
# Exemple :
data(anscombe)
names(anscombe)
attach(anscombe)
modele=lm(y1~x1)
modele
summary(modele)
```

#### 3. Vous pouvez afficher vos sorties graphique

```{r,} 
plot(x1,y1,xlab="Legend x", ylab="Legend y", main="Titre", pch=19)
abline(modele,col=1)
```

#### 4. Vous pouvez écrire des équations mathématiques

$$
\forall i=1,\dots,n\quad Y_i = \beta + \alpha_1,x_{i,1} + \alpha_2,x_{i,2} +  
+ \alpha_3,x_{i,3} + \alpha_4,x_{i,4} + \alpha_5,x_{i,5} + \alpha_6,x_{i,6} +
+ \alpha_7,x_{i,7} + \alpha_8,x_{i,8} +
\varepsilon_i
$$

où

+ $\alpha_1, \ldots, \alpha_{p}$ et $\beta$ sont les paramètres d'espérance du modèle.
+ $\varepsilon_i$ sont les erreurs résiduelles aléatoires supposées iid de loi ${\cal N}(0,\sigma^2)$.
+ $\sigma^2$ est le paramètre de variance.

#### 5. Vous pouvez également faire des tableaux 

Critère | Type de selection | Variables explicatives | $R^2$ | $R^2_{ajus}$ | 
------- | ----------------- | ---------------------- | ----- | ------------ |
AIC     | Descendante       | $x_1,x_3,x_5$          | 0.67  | 0.54         |
BIC     | Ascendante        | $x_1,x_5$              | 0.61  | 0.49         |



